---
output: html_document
editor_options: 
  chunk_output_type: console
---

# Statistical Analysis

## Packages 

```{r}
options(scipen = 999) ##notacion cientifica apagada
rm(list=ls())
#install.packages("tidyverse")
library(tidyverse)
#install.packages("skimr")
library(skimr)
#install.packages("here")
library(here)
```

## Dataset

```{r}
data <- read_csv("DATA_Final.csv")
dim(data)
head(data)

skim(data) # Análisis descriptivo de todas las variables.
```

## Distribution

```{r}
# Fluoride
summary(data$LBDPFL)
hist(data$LBDPFL, breaks=15)
hist(log(data$LBDPFL), breaks=15)

# Lead
summary(data$LBXBPB)
hist(data$LBXBPB, breaks=15)
hist(log(data$LBXBPB), breaks=15)

# Manganese
summary(data$LBXBMN)
hist(data$LBXBMN, breaks=15)
hist(log(data$LBXBMN), breaks=15)

# Selenium
summary(data$LBXBSE)
hist(data$LBXBSE, breaks=15)
hist(log(data$LBXBSE), breaks=15)
```

## Exposure variables

```{r, results='hide', message=FALSE}
data$ln_LBDPFL <- log(data$LBDPFL)
data$ln_LBXBMN <- log(data$LBXBMN)
data$ln_LBXBPB <- log(data$LBXBPB)
data$ln_LBXBSE <- log(data$LBXBSE)
data$ln_LBXTHG <- log(data$LBXTHG)

data$iqrFl <- data$ln_LBDPFL/IQR(data$ln_LBDPFL)
data$iqrMn <- data$ln_LBXBMN/IQR(data$ln_LBXBMN)
data$iqrPb <- data$ln_LBXBPB/IQR(data$ln_LBXBPB)
data$iqrSe <- data$ln_LBXBSE/IQR(data$ln_LBXBSE)
data$iqrHg <- data$ln_LBXTHG/IQR(data$ln_LBXTHG)
```

## Confounding factors
```{r, results='hide', message=FALSE}
data$calories <- as.numeric(data$DR1TKCAL)
data$INDFMPIR_ts <- factor(data$INDFMPIR_ts_low.high, levels = c("low","high"), labels = c(0,1))
data$sex <- factor(data$RIAGENDR, levels = c(1,2), labels = c(0,1))
data$age <- as.numeric(data$RIDAGEYR)
data$race <- as.factor(data$RIDRETH3.ts)
data$smoke <- as.factor(data$SMD460.ts)
data$activityscore
data$sch_activityscore
```

## Fluoride (F)

### BMI - *Body Mass Index*
```{r, results='hide', message=FALSE}
Multiple_BMXBMI <- lm(data$BMXBMI ~  (data$iqrFl) + 
                        (data$iqrMn) +
                        (data$iqrPb) +
                        (data$iqrSe) + 
                        (data$iqrHg) + 
                        data$calories + data$sex + data$INDFMPIR_ts + data$age + data$smoke + data$activityscore + data$sch_activityscore + data$race) 
summary(Multiple_BMXBMI)
round(confint(Multiple_BMXBMI), 2)
```

### Standing height
```{r, results='hide', message=FALSE}
Multiple_BMXHT <- lm(data$BMXHT ~  (data$iqrFl) + 
                       (data$iqrMn) +
                       (data$iqrPb) +
                       (data$iqrSe) + 
                       (data$iqrHg) +
                       data$calories + data$sex + data$INDFMPIR_ts + data$age + data$smoke + data$activityscore + data$sch_activityscore + data$race) 
summary(Multiple_BMXHT)
round(confint(Multiple_BMXHT),2)
```

### Waist circumference

```{r, results='hide', message=FALSE}
Multiple_BMXWAIST <- lm(data$BMXWAIST ~  (data$iqrFl) + 
                          (data$iqrMn) +
                          (data$iqrPb) +
                          (data$iqrSe) + 
                          (data$iqrHg) +
                          data$calories + data$sex + data$INDFMPIR_ts + data$age + data$smoke + data$activityscore + data$sch_activityscore + data$race) 
summary(Multiple_BMXWAIST)
round(confint(Multiple_BMXWAIST), 2)
```

### Upper arm length
```{r, results='hide', message=FALSE}
Multiple_BMXARML <- lm(data$BMXARML ~  (data$iqrFl) + 
                         (data$iqrMn) +
                         (data$iqrPb) +
                         (data$iqrSe) + 
                         (data$iqrHg) +
                         data$calories + data$sex + data$INDFMPIR_ts + data$age + data$smoke + data$activityscore + data$sch_activityscore + data$race) 
summary(Multiple_BMXARML)
round(confint(Multiple_BMXARML), 2)
```

## Lead (Pb)

### Body Mass Index

```{r, results='hide', message=FALSE}
Multiple_BMXBMI <- lm(data$BMXBMI ~  (data$iqrPb) + 
                        (data$iqrMn) +
                        (data$iqrFl) +
                        (data$iqrSe) + 
                        (data$iqrHg) + 
                        data$calories + data$sex + data$INDFMPIR_ts + data$age + data$smoke + data$activityscore + data$sch_activityscore + data$race) 
summary(Multiple_BMXBMI)
round(confint(Multiple_BMXBMI), 2)
```

### Standing height

```{r, results='hide', message=FALSE}
Multiple_BMXHT <- lm(data$BMXHT ~  (data$iqrPb) + 
                       (data$iqrMn) +
                       (data$iqrFl) +
                       (data$iqrSe) + 
                       (data$iqrHg) +
                       data$calories + data$sex + data$INDFMPIR_ts + data$age + data$smoke + data$activityscore + data$sch_activityscore + data$race) 
summary(Multiple_BMXHT)
round(confint(Multiple_BMXHT),2)
```

### Waist circumference
```{r, results='hide', message=FALSE}
Multiple_BMXWAIST <- lm(data$BMXWAIST ~  (data$iqrPb) + 
                          (data$iqrMn) +
                          (data$iqrFl) +
                          (data$iqrSe) + 
                          (data$iqrHg) +
                          data$calories + data$sex + data$INDFMPIR_ts + data$age + data$smoke + data$activityscore + data$sch_activityscore + data$race) 
summary(Multiple_BMXWAIST)
round(confint(Multiple_BMXWAIST), 2)
```

### Upper arm length

```{r, results='hide', message=FALSE}
Multiple_BMXARML <- lm(data$BMXARML ~  (data$iqrPb) + 
                         (data$iqrMn) +
                         (data$iqrFl) +
                         (data$iqrSe) + 
                         (data$iqrHg) +
                         data$calories + data$sex + data$INDFMPIR_ts + data$age + data$smoke + data$activityscore + data$sch_activityscore + data$race) 
summary(Multiple_BMXARML)
round(confint(Multiple_BMXARML), 2)
```

## Manganese (Mn)

### Body Mass Index
```{r, results='hide', message=FALSE}
Multiple_BMXBMI <- lm(data$BMXBMI ~  (data$iqrMn) + 
                        (data$iqrPb) +
                        (data$iqrFl) +
                        (data$iqrSe) + 
                        (data$iqrHg) + 
                        data$calories + data$sex + data$INDFMPIR_ts + data$age + data$smoke + data$activityscore + data$sch_activityscore + data$race) 
summary(Multiple_BMXBMI)
round(confint(Multiple_BMXBMI), 2)
```

### Standing height
```{r, results='hide', message=FALSE}
Multiple_BMXHT <- lm(data$BMXHT ~  (data$iqrMn) + 
                       (data$iqrPb) +
                       (data$iqrFl) +
                       (data$iqrSe) + 
                       (data$iqrHg) +
                       data$calories + data$sex + data$INDFMPIR_ts + data$age + data$smoke + data$activityscore + data$sch_activityscore + data$race) 
summary(Multiple_BMXHT)
round(confint(Multiple_BMXHT),2)
```

### Waist circumference
```{r, results='hide', message=FALSE}
Multiple_BMXWAIST <- lm(data$BMXWAIST ~  (data$iqrMn) + 
                          (data$iqrPb) +
                          (data$iqrFl) +
                          (data$iqrSe) + 
                          (data$iqrHg) +
                          data$calories + data$sex + data$INDFMPIR_ts + data$age + data$smoke + data$activityscore + data$sch_activityscore + data$race) 
summary(Multiple_BMXWAIST)
round(confint(Multiple_BMXWAIST), 2)
```

### Upper arm length
```{r, results='hide', message=FALSE}
Multiple_BMXARML <- lm(data$BMXARML ~  (data$iqrMn) + 
                         (data$iqrPb) +
                         (data$iqrFl) +
                         (data$iqrSe) + 
                         (data$iqrHg) +
                         data$calories + data$sex + data$INDFMPIR_ts + data$age + data$smoke + data$activityscore + data$sch_activityscore + data$race) 
summary(Multiple_BMXARML)
round(confint(Multiple_BMXARML), 2)
```

## Mercury (Hg)

### Body Mass Index

```{r, results='hide', message=FALSE}
Multiple_BMXBMI <- lm(data$BMXBMI ~  (data$iqrHg) + 
                        (data$iqrPb) +
                        (data$iqrFl) +
                        (data$iqrSe) + 
                        (data$iqrMn) + 
                        data$calories + data$sex + data$INDFMPIR_ts + data$age + data$smoke + data$activityscore + data$sch_activityscore + data$race) 
summary(Multiple_BMXBMI)
round(confint(Multiple_BMXBMI), 2)
```

### Standing height

```{r, results='hide', message=FALSE}
Multiple_BMXHT <- lm(data$BMXHT ~  (data$iqrHg) + 
                       (data$iqrPb) +
                       (data$iqrFl) +
                       (data$iqrSe) + 
                       (data$iqrMn) +
                       data$calories + data$sex + data$INDFMPIR_ts + data$age + data$smoke + data$activityscore + data$sch_activityscore + data$race) 
summary(Multiple_BMXHT)
round(confint(Multiple_BMXHT),2)
```

### Waist circumference

```{r, results='hide', message=FALSE}
Multiple_BMXWAIST <- lm(data$BMXWAIST ~  (data$iqrHg) + 
                          (data$iqrPb) +
                          (data$iqrFl) +
                          (data$iqrSe) + 
                          (data$iqrMn) +
                          data$calories + data$sex + data$INDFMPIR_ts + data$age + data$smoke + data$activityscore + data$sch_activityscore + data$race) 
summary(Multiple_BMXWAIST)
round(confint(Multiple_BMXWAIST), 2)
```

### Upper arm length
```{r, results='hide', message=FALSE}
Multiple_BMXARML <- lm(data$BMXARML ~  (data$iqrHg) + 
                         (data$iqrPb) +
                         (data$iqrFl) +
                         (data$iqrSe) + 
                         (data$iqrMn) +
                         data$calories + data$sex + data$INDFMPIR_ts + data$age + data$smoke + data$activityscore + data$sch_activityscore + data$race) 
summary(Multiple_BMXARML)
round(confint(Multiple_BMXARML), 2)
```

## Selenium (Se)

### Body Mass Index
```{r, results='hide', message=FALSE}
Multiple_BMXBMI <- lm(data$BMXBMI ~  (data$iqrSe) + 
                        (data$iqrPb) +
                        (data$iqrFl) +
                        (data$iqrHg) + 
                        (data$iqrMn) + 
                        data$calories + data$sex + data$INDFMPIR_ts + data$age + data$smoke + data$activityscore + data$sch_activityscore + data$race) 
summary(Multiple_BMXBMI)
round(confint(Multiple_BMXBMI), 2)
```

### standing height
```{r, results='hide', message=FALSE}
Multiple_BMXHT <- lm(data$BMXHT ~  (data$iqrSe) + 
                       (data$iqrPb) +
                       (data$iqrFl) +
                       (data$iqrHg) + 
                       (data$iqrMn) +
                       data$calories + data$sex + data$INDFMPIR_ts + data$age + data$smoke + data$activityscore + data$sch_activityscore + data$race) 
summary(Multiple_BMXHT)
round(confint(Multiple_BMXHT),2)
```

### Waist circumference
```{r, results='hide', message=FALSE}
Multiple_BMXWAIST <- lm(data$BMXWAIST ~  (data$iqrSe) + 
                          (data$iqrPb) +
                          (data$iqrFl) +
                          (data$iqrHg) + 
                          (data$iqrMn) +
                          data$calories + data$sex + data$INDFMPIR_ts + data$age + data$smoke + data$activityscore + data$sch_activityscore + data$race) 
summary(Multiple_BMXWAIST)
round(confint(Multiple_BMXWAIST), 2)
```

### Upper arm length
```{r, results='hide', message=FALSE}
Multiple_BMXARML <- lm(data$BMXARML ~  (data$iqrSe) + 
                         (data$iqrPb) +
                         (data$iqrFl) +
                         (data$iqrHg) + 
                         (data$iqrMn) +
                         data$calories + data$sex + data$INDFMPIR_ts + data$age + data$smoke + data$activityscore + data$sch_activityscore + data$race) 
summary(Multiple_BMXARML)
round(confint(Multiple_BMXARML), 2)
```
